### 2020-02-25 CalNotes.tex
Figures fix
Modifications after discussion with Assyr

\documentclass[10pt]{article}
%\documentclass{siamart1116}

@@

Let us consider the filtering kernel \eqref{eq:filter} with $\beta = 1$, i.e.,
\begin{equation}\label{eq:filter_beta1}
	k(r) = \frac{1}{\delta} e^{-r/\delta}.
\end{equation}
In this case, Leibniz integral rule yields the equality
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t k'(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t,
\end{equation}
which can be interpreted as an ordinary differential equation for $Z_t^\epl$ driven by the stochastic signal $X^\epl$. Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain the system of two one-dimensional SDEs
\begin{equation}
\label{eq:systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
The first ingredient for verifying the ergodic properties of the two-dimensional process $(X^\epl, Z^\epl)^\top \defeq ((X^\epl_t, Z^\epl_t)^\top, 0 \leq t \leq T)$ is verifying that the measure induced by the stochastic process admits a smooth density with respect to the Lebesgue measure. Since noise is present only on the first component, this is a consequence of the theory of hypo-ellipticity, as summarized in the following Lemma.

@@

\subsection{A second filtering-based estimator}\label{sec:FilterMLE_2}

In this section we consider a modification to the estimator \eqref{eq:AHatMixed}, which is defined as
\begin{equation}\label{eq:AHatMixedTilde}
	\widetilde A^\epl_k(T) = -M^{-1}\widetilde h,
\end{equation}
where the matrix $M$ is taken from \eqref{eq:MandH} and the vector $\widetilde{h}$ is taken from \eqref{eq:MandHTilde}. For this estimator we will be able to prove a result of unbiasedness under a condition on the filtering parameter $\delta$. Let us remark that for $\widehat A^\epl_k(T)$ we do not need any condition on this parameter. Nevertheless, this second estimator is the minimizer of a likelihood function unlike the first, and will therefore be used in the Bayesian framework. Let us start with an estimate concerning the distance between $X^\epl$ and $Z^\epl$.

\begin{theorem}\label{thm:mainTheoremTilde} Let the assumptions of Theorem \ref{thm:mainTheorem} hold. Then, if $\delta$ in \eqref{eq:filter} satisfies $\delta = \delta(\epl)$ with $\delta(\epl) \to 0$ for $\epl \to 0$, it holds
	\begin{equation}
		\lim_{\epl \to 0} \lim_{T \to \infty} \widetilde A^\epl_k(T) = A, \quad \text{a.s.},
	\end{equation} 
	for $\widetilde A^\epl_k(T)$ defined in \eqref{eq:AHatMixedTilde}.
\end{theorem}
\begin{proof} We first consider the difference between the two estimators $\widetilde A^\epl_k(T)$ and $\widehat A^\epl_k(T)$. In particular, the ergodic theorem and an algebraic equality imply
	\begin{equation}
	\begin{aligned}
		\lim_{T \to \infty} \left(\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)\right) &= \left(\mathcal M^{-1} - \widetilde{\mathcal M}^{-1}\right) \lim_{T \to \infty}\widetilde h \\
		&= \mathcal M^{-1}\left(\mathcal M - \widetilde{\mathcal M}\right)  \widetilde{\mathcal M}^{-1} \lim_{T \to \infty} \widetilde h\\
		&= \mathcal M^{-1}\left(\mathcal M - \widetilde{\mathcal M}\right)\lim_{T \to \infty} \widehat A^\epl_k(T),
	\end{aligned}
	\end{equation}
	almost surely. Therefore, due to Assumption \ref{as:regularity} which allows controlling the norm of $\mathcal M^{-1}$ and due to Lemma \ref{lem:distanceMandTildeM} we have for a constant $C > 0$
	\begin{equation}\label{eq:DistATildeAHat}
		\lim_{T \to \infty} \norm{\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)}_2 \leq C \left(\epl + \delta^{1/2} \right).
	\end{equation}
	Let us remark that $\widehat A^\epl_k(T)$ has a bounded norm for $\epl$ sufficiently small due to Theorem \ref{thm:mainTheorem}. Now, the triangle inequality yields
	\begin{equation}
		\norm{\widetilde A^\epl_k(T) - A}_2 \leq \norm{\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)}_2 + \norm{\widehat A^\epl_k(T) - A}_2.
	\end{equation}
	Therefore, due to Theorem \ref{thm:mainTheorem}, the inequality \eqref{eq:DistATildeAHat} and since $\delta \to 0$ for $\epl \to 0$, the desired result holds.
\end{proof}

@@

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. In particular, we employ the same strategy that led to the estimator $\widetilde A^\epl_k(T)$ presented in Section \ref{sec:FilterMLE_2}. Therefore, we modify the likelihood function as
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
	\widetilde I(X^\epl \mid A) =\int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t,
\end{equation}
so that the estimator $\widetilde A^\epl_k(T)$ introduced in \eqref{eq:AHatMixedTilde} is the MLE based on the new likelihood $\widetilde p(X^\epl \mid A)$. With this likelihood function, we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + \frac{T}{2\Sigma} M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - \frac{T}{2\Sigma} \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$. The following result now holds.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}

%In this section we show how the filtering approach can be employed in a Bayesian framework. As in Section \ref{sec:Bayesian} let us consider a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$. Then, the density $p^\epl(A \mid X)$ of the posterior distribution $\mu^\epl(A \mid X^\epl)$ given by multiscale data satisfies
%\begin{equation}
%	p^\epl(A \mid X) = \frac1{Z^\epl} \, p^\epl(X \mid A) \, p_0(A),
%\end{equation}
%where $Z^\epl$ is the normalization term and where the likelihood function is given by
%\begin{equation}
%	p^\epl(X \mid A) = \exp\left(-\frac{I^\epl(X\mid A)}{2\Sigma} \right), 
%\end{equation}
%with
%\begin{equation}
%	I^\epl(X\mid A) = \int_0^T A \cdot V'(X_t^\epl) \dd X_t^\epl + \frac12 \int_0^T \left( A \cdot V'(X_t^\epl) \right)^2 \dd t.
%\end{equation}
%Let us consider the likelihood function $\widetilde p^\epl(X \mid A)$ modified to include the filtered data $Z^\epl$ given by
%\begin{equation}
%	\widetilde p^\epl(X \mid A) = \exp\left(-\frac{\widetilde I^\epl(X\mid A)}{2\Sigma} \right), 
%\end{equation}
%with
%\begin{equation}
%	\widetilde I^\epl(X\mid A) = \int_0^T A \cdot V'(Z_t^\epl) \dd X_t^\epl + \frac12 \int_0^T \left( A \cdot V'(X_t^\epl) \right) \left( A \cdot V'(Z_t^\epl) \right) \dd t.
%\end{equation}
%Moreover, let us consider the modified posterior distribution $\widetilde \mu^\epl(A \mid X)$ with density
%\begin{equation}
%	\widetilde p^\epl(A \mid X) = \frac1{\widetilde Z^\epl} \, \widetilde p^\epl(X \mid A) \, p_0(A),
%\end{equation}
%Let us remark that employing \eqref{eq:MandHTilde} we can write 
%\begin{equation}
%	\log \widetilde p^\epl(A \mid X) = -\log \widetilde Z^\epl - T A \cdot \widetilde h - \frac{T}{2} A \cdot \widetilde M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
%\end{equation}
%With respect to Section \ref{sec:Bayesian}, the matrix $\widetilde M$ is not symmetric. Nonetheless, we remark that
%\begin{equation}
%	A \cdot \widetilde M A = A \cdot \widetilde M_S A,
%\end{equation}
%where $\widetilde M_S$ is the symmetric part of $\widetilde M$, i.e.,
%\begin{equation}
%	\widetilde M_S = \frac12 \left(\widetilde M + \widetilde M^\top\right).
%\end{equation}
%Hence, we can rewrite 
%\begin{equation}	
%	\log \widetilde p^\epl(A \mid X) = -\log \widetilde Z^\epl - T A \cdot \widetilde h - \frac{T}{2} A \cdot \widetilde M_S A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
%\end{equation}
%This shows that the posterior $\widetilde \mu^\epl$ is a Gaussian $\mathcal N(\widetilde m_T, \widetilde C_T)$ where the covariance and mean can be obtained as 
%\begin{equation}
%\begin{aligned}
%	\widetilde C_T^{-1} &= C_0^{-1} + T \widetilde M, \\
%	\widetilde C_T^{-1} \widetilde m_T &= C_0^{-1}A_0 - T \widetilde h. 
%\end{aligned}
%\end{equation}
%
%In the following theorem we show that the posterior we obtain 
%\begin{theorem} Under
%\end{theorem}

@@

@@

\subsection{Multi-dimensional parameter}

\begin{table}[t]
	\centering
	\begin{tabular}{ccccc}
		\toprule
		Coefficient & Truth & No preprocessing & Filtering & Subsampling  \\ 
		\midrule
		$A_1$ & -0.62 & -0.92 & -0.70 & -0.59\\
		$A_2$ & -0.31 & -0.70 & -0.27 & 0.05 \\
		$A_3$ & 0.31  & 0.55  & 0.31  & 0.14 \\
		$A_4$ & 0.62  & 1.22  & 0.57  & 0.13 \\
		\bottomrule
	\end{tabular}
	\caption{Numerical results for the four-dimensional parameter rounded to the second significant figure.}
	\label{tab:KLStyle}
\end{table}

@@

>>>>

\documentclass[10pt]{article}

@@

Let us consider the filtering kernel \eqref{eq:filter} with $\beta = 1$, i.e.,
\begin{equation}\label{eq:filter_beta1}
	k(r) = \frac{1}{\delta} e^{-r/\delta}.
\end{equation}
In this case, Leibniz integral rule yields the equality
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t k'(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t,
\end{equation}
which can be interpreted as an ordinary differential equation for $Z_t^\epl$ driven by the stochastic signal $X^\epl$. Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain the system of two one-dimensional SDEs
\begin{equation}
\label{eq:systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) \dd t+ \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
The first ingredient for verifying the ergodic properties of the two-dimensional process $(X^\epl, Z^\epl)^\top \defeq ((X^\epl_t, Z^\epl_t)^\top, 0 \leq t \leq T)$ is verifying that the measure induced by the stochastic process admits a smooth density with respect to the Lebesgue measure. Since noise is present only on the first component, this is a consequence of the theory of hypo-ellipticity, as summarized in the following Lemma.

@@

@@

In this section, we present how to correct the faulty behaviour highlighted by Theorem \ref{thm:BayesianBias} with filtering techniques. Employing the filter in the same way as for the MLE, we can modify the likelihood function and obtain 
\begin{equation}
	\widehat p(X^\epl \mid A) = \exp\left(-\frac{\widehat I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
\begin{aligned}
	\widehat I(X^\epl \mid A) & = \int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left(A \cdot V'(X^\epl_t)\right)\left(A \cdot V'(Z^\epl_t)\right) \dd t \\
	&= \widetilde h \cdot A + \frac12 A \cdot \widetilde M A.
\end{aligned}
\end{equation}
Let us remark that $\widetilde M$ is not symmetric. Nevertheless, we have that
\begin{equation}
	A \cdot \widetilde M A = A \cdot \widetilde M_S A,
\end{equation}
where $\widetilde M_S$ is the symmetric part of $\widetilde M$. Repeating the same reasoning as above, one would therefore obtain a modified posterior $\widehat \mu_{T, \epl} = \mathcal N(\widehat m_{T, \epl}, \widehat C_{T, \epl})$, whose covariance matrix and mean would be given by
\begin{equation}
\begin{aligned}
	\widehat C_{T, \epl}^{-1} &= C_0^{-1} + \frac{T}{2\Sigma} \widetilde M_S, \\
	\widehat C_{T,\epl}^{-1}\widehat m_{T,\epl} &= C_0^{-1}A_0 - \frac{T}{2\Sigma} \widetilde h. 
\end{aligned}	
\end{equation}
There is no guarantee under Assumption \ref{as:regularity} that $\widehat C_{T, \epl}$ is positive define and therefore a covariance matrix. Hence, in this Bayesian setting, we consider a different modification of the likelihood where we perturb only the vector $h$, i.e.,
\begin{equation}
	\widetilde p(X^\epl \mid A) = \exp\left(-\frac{\widetilde I(X^\epl\mid A)}{2\Sigma} \right), 
\end{equation}
where 
\begin{equation}
\begin{aligned}
	\widetilde I(X^\epl \mid A) & = \int_0^T A \cdot V'(Z^\epl_t) \dd X^\epl_t + \frac12 \int_0^T \left(A \cdot V'(X^\epl_t)\right)^2 \dd t \\
	&= \widetilde h \cdot A + \frac12 A \cdot M A.
\end{aligned}
\end{equation}
so that we obtain the modified posterior $\widetilde \mu_{T,\epl} = \mathcal N(\widetilde m_{T, \epl}, C_{T, \epl})$, whose parameters are given by
\begin{equation}
\begin{aligned}
	C_{T, \epl}^{-1} &= C_0^{-1} + \frac{T}{2\Sigma} M, \\
	C_{T,\epl}^{-1}\widetilde m_{T,\epl} &= C_0^{-1}A_0 - \frac{T}{2\Sigma} \widetilde h. 
\end{aligned}	
\end{equation}
Let us remark that the posterior $\widetilde \mu_{T,\epl}$ has the same covariance as $\mu_{T,\epl}$ and that therefore it is indeed a valid Gaussian posterior distribution. Nevertheless, in order to employ the tool of convergence introduced in Theorem \ref{thm:BayesianBias}, we need to study the properties of the MLE based on the likelihood $\widetilde p(X^\epl \mid A)$, i.e., the quantity
\begin{equation}\label{eq:AHatMixedTilde}
	\widetilde A^\epl_k(T) = - M^{-1} \widetilde h.
\end{equation}
The following theorem guarantees the unbiasedness of this estimator under a condition on the parameter $\delta$ of the filter, and we refer to the Appendix for its proof.
\begin{theorem}\label{thm:mainTheoremTilde} Let the assumptions of Theorem \ref{thm:mainTheorem} hold. Then, if $\delta$ in \eqref{eq:filter} satisfies $\delta = \delta(\epl)$ with $\delta(\epl) \to 0$ for $\epl \to 0$, it holds
	\begin{equation}
	\lim_{\epl \to 0} \lim_{T \to \infty} \widetilde A^\epl_k(T) = A, \quad \text{a.s.},
	\end{equation} 
	for $\widetilde A^\epl_k(T)$ defined in \eqref{eq:AHatMixedTilde}.
\end{theorem}

The following result gives the convergence of the modified posterior distribution $\widetilde \mu_{T, \epl}$ to the drift coefficient of the homogenized equation.
\begin{theorem} Let the Assumptions of Theorem \ref{thm:mainTheoremTilde} hold. Then, the modified posterior measure $\widetilde\mu_{T,\epl} = \mathcal N(\widetilde m_{T,\epl}, C_{T,\epl})$ satisfies
	\begin{equation}
	\lim_{\epl \to 0}\lim_{T \to \infty} \widetilde \mu_{T,\epl}(\cdot) = \delta(\cdot - A), \quad \text{a.s.},
	\end{equation}
	weakly in the space of probability measures on $\R^N$, where $\delta$ is the Dirac measure on $\R^N$.
\end{theorem}
\begin{proof} The proof follows from the proof of Theorem \ref{thm:BayesianBias} and from Theorem \ref{thm:mainTheoremTilde}.
\end{proof}

@@

\subsection{Multi-dimensional drift coefficient}

\begin{table}[ht!]
	\centering
	\begin{tabular}{cccccc}
		\toprule
		Coefficient & Multiscale & Homogenized & No preprocessing  & Filtering  		  & Subsampling  			 \\ 
		& $\alpha$ & $A$   & $\widehat A^\epl$ & $\widehat A^\epl_k$ & $\widehat A^\epl_\delta$ \\
		\midrule
		$1$ &-1   & -0.62 & -0.92 & -0.70 & -0.59\\
		$2$ &-0.5 & -0.31 & -0.70 & -0.27 & 0.05 \\
		$3$ & 0.5 & 0.31  & 0.55  & 0.31  & 0.14 \\
		$4$ & 1   & 0.62  & 1.22  & 0.57  & 0.13 \\
		\bottomrule
	\end{tabular}
	\caption{Numerical results for the four-dimensional parameter rounded to the second significant digit.}
	\label{tab:KLStyle}
\end{table}

@@

@@

\begin{appendices}

\section{}
Let us consider the estimator defined in \eqref{eq:AHatMixedTilde}, that is
\begin{equation}
	\widetilde A^\epl_k(T) = -M^{-1}\widetilde h,
\end{equation}
where the matrix $M$ is taken from \eqref{eq:MandH} and the vector $\widetilde{h}$ is taken from \eqref{eq:MandHTilde}. We prove here Theorem \ref{thm:mainTheoremTilde}, i.e., the unbiasedness of this estimator, for which it is necessary to prove two technical lemmas. Let us start with an estimate concerning the distance between $X^\epl$ and $Z^\epl$.

\begin{proof}[Proof of Theorem \ref{thm:mainTheoremTilde}] We first consider the difference between the two estimators $\widetilde A^\epl_k(T)$ and $\widehat A^\epl_k(T)$. In particular, the ergodic theorem and an algebraic equality imply
	\begin{equation}
	\begin{aligned}
	\lim_{T \to \infty} \left(\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)\right) &= \left(\mathcal M^{-1} - \widetilde{\mathcal M}^{-1}\right) \lim_{T \to \infty}\widetilde h \\
	&= \mathcal M^{-1}\left(\mathcal M - \widetilde{\mathcal M}\right)  \widetilde{\mathcal M}^{-1} \lim_{T \to \infty} \widetilde h\\
	&= \mathcal M^{-1}\left(\mathcal M - \widetilde{\mathcal M}\right)\lim_{T \to \infty} \widehat A^\epl_k(T),
	\end{aligned}
	\end{equation}
	almost surely. Therefore, due to Assumption \ref{as:regularity} which allows controlling the norm of $\mathcal M^{-1}$ and due to Lemma \ref{lem:distanceMandTildeM} we have for a constant $C > 0$
	\begin{equation}\label{eq:DistATildeAHat}
	\lim_{T \to \infty} \norm{\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)}_2 \leq C \left(\epl + \delta^{1/2} \right).
	\end{equation}
	Let us remark that $\widehat A^\epl_k(T)$ has a bounded norm for $\epl$ sufficiently small due to Theorem \ref{thm:mainTheorem}. Now, the triangle inequality yields
	\begin{equation}
	\norm{\widetilde A^\epl_k(T) - A}_2 \leq \norm{\widetilde A^\epl_k(T) - \widehat A^\epl_k(T)}_2 + \norm{\widehat A^\epl_k(T) - A}_2.
	\end{equation}
	Therefore, due to Theorem \ref{thm:mainTheorem}, the inequality \eqref{eq:DistATildeAHat} and since $\delta \to 0$ for $\epl \to 0$, the desired result holds.
\end{proof}
\end{appendices}
